---
title: "My Spotify Data - 2022"
author: "Sam Freis"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
cat("\014") 

library(jsonlite)
library(lubridate)
library(gghighlight)
library(spotifyr)
library(tidyverse)
library(knitr)
library(ggplot2)
library(plotly)
```

## Analyzing and visualizing my own Spotify history from January 2022 - January 2023 

#### Data Import ####
```{r}
# read in stream history
stream <- fromJSON("StreamingHistory0.json", flatten = TRUE)

```

#### Brief Data Examination of Data Structure ####
```{r, warning=FALSE}
# look at var names and info
names(stream)
str(stream)
head(stream)

# add date and time
spotify <- stream %>% 
  as_tibble() %>% 
  mutate_at("endTime", ymd_hm) %>% 
  mutate(endTime = endTime - hours(6)) %>% 
  mutate(date = floor_date(endTime, "day") %>% as_date, seconds = msPlayed / 1000, minutes = seconds / 60)
```

## Research Questions: 
### 1. Which artists did I listen to most?
```{r, warning=FALSE}
p1 <- spotify %>% 
  filter(date >= "2022-01-01") %>% 
  group_by(artistName) %>% 
  summarize(minutesListened = sum(minutes)) %>% 
  filter(minutesListened >= 180) %>%
  ggplot(aes(x = artistName, y = minutesListened)) + 
  geom_col(aes(fill = minutesListened)) +
  scale_fill_gradient(low = "grey", high = "darkblue") + 
  labs(x= "Artist", y= "Minutes Listened") + 
  ggtitle("Top Artists by Minutes Listened") +
  theme_minimal()+
  theme(legend.position = "bottom", axis.text.x = element_text(angle=60, size=7, hjust = 1))
p1

# activity by artist 
p2 <- spotify %>% 
  group_by(artistName, date = floor_date(date, "month")) %>% 
  summarize(hours = sum(minutes) / 60) %>% 
  ggplot(aes(x = date, y = hours, group = artistName)) + 
  labs(x= "Date", y= "Hours of Listen Time") + 
  ggtitle("Listening Activity by Artist Over Time") +
  geom_line() + 
  gghighlight(artistName == "St. Vincent" || artistName == "Hovvdy")+
  theme_minimal()
p2
```

My top artists in this time frame were both Texans, of course. I listened to Hovvdy and St. Vincent for 750+ minutes last year. My Hovvdy listening peaked right after they released a new album and right around when I saw them live. My St. Vincent listening was high at the start of 2022, around when I got to see her in concert, and then my listening took off again in late 2022. I tend to listen to music I know well when working on analyses and writing. So, St. Vincent's discography has been playing on repeat while I worked on projects with deadlines at the end of 2022. Also, shout out to the Quantitude Pod for making the 3+ hours of listening time cut. 

### 2. Which songs did I listen to most?
```{r, warning=FALSE}
p3<-spotify %>% 
  filter(date >= "2022-01-01") %>% 
  group_by(trackName) %>% 
  summarize(minutesListened = sum(minutes)) %>% 
  filter(minutesListened >= 100) %>%
  ggplot(aes(x = trackName, y = minutesListened)) + 
  geom_col(aes(fill = minutesListened)) +
  scale_fill_gradient(low = "grey", high = "darkorange3") + 
  labs(x= "Track", y= "Minutes Listened") + 
  ggtitle("Top Tracks by Minutes Listened") +
  theme_minimal()+
  theme(legend.position = "bottom", axis.text.x = element_text(angle=60, size=7, hjust = 1))
p3
```

Songs I listened to for over 100 minutes in this time frame. My top songs were concussion by girlhouse and Speeding 72 by Momma. 

### 3. Looking at time-related data
```{r, warning=FALSE}
# activity per week and hour
p4 <- spotify %>% 
  filter(date >= "2022-01-01") %>% 
  group_by(date) %>% 
  group_by(date = floor_date(date, "week")) %>%
  summarize(hours = sum(minutes) / 60) %>% 
  arrange(date) %>% 
  ggplot(aes(x = date, y = hours)) + 
  geom_col(aes(fill = hours)) +
  scale_fill_gradient(low = "grey", high = "darkgreen") + 
  labs(x= "Date", y= "Hours of Listen Time") + 
  ggtitle("Listening Activity per Week")+
  theme_minimal()
p4

# activity by time of day 
day <- spotify %>% 
  filter(date >= "2022-01-01") %>% 
  group_by(date, hour = hour(endTime), weekday = wday(date, label = TRUE))%>% 
  summarize(minutesListened = sum(minutes))

p5<-day %>% 
  ggplot(aes(x = hour, y = minutesListened, group = date)) + 
  geom_col(fill = "darkgreen") +
  labs(x= "Time of Day (24hr)", y= "Minutes of Listen Time")+
  scale_x_continuous(breaks = seq(0, 24, by = 2))+
  theme_minimal()
p5

# activity by day
p6 <-day %>% 
  group_by(weekday, hour) %>% 
  summarize(minutes = sum(minutesListened)) %>% 
  ggplot(aes(x = weekday, hour, fill = minutes)) + 
  geom_tile() + 
  scale_fill_gradient(low = "darkblue", high = "red") +
  labs(x= "Weekday", y= "Time of Day (24hr)")+
  scale_y_continuous(breaks = seq(0, 24, by = 2))+
  theme_minimal()+
  theme(legend.position = "bottom", axis.text.x = element_text(angle=60, size=7, hjust = 1))
p6

# activity weekdays v weekends
p7 <- day %>% 
  mutate(day_type = if_else(weekday %in% c("Sat", "Sun"), "weekend", "weekday")) %>% 
  group_by(day_type, hour) %>% 
  summarize(minutes = sum(minutesListened)) %>% 
  ggplot(aes(x = day_type, y = minutes, color = day_type)) + 
  geom_boxplot()+
  labs(x= "Day Type", y= "Minutes of Listen Time") + 
theme_minimal()+
  theme(axis.text.x = element_text(size=10, hjust = 1))
p7
```

I listened to a lot music in between April of 2022 and August of 2022. I drove to Texas from Colorado and back that summer, so these trends make sense. My listening time typically peaks on weekday mornings.  
